---
title: "report"
author: "PLR"
date: "2024-01-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, cache = FALSE)
```

## Setup

Load required packages first

```{r load packages}
# "tibble","tidyverse","here"
library(tibble)
library(tidyverse)
library(here)
library(targets)
```

## Load Targets functions

We source required Target functions to create the output elements for this report:

```{r}
source(here("R","study_functions.R"))
```

## Create objects from functions

This adhoc. source_all() function can be used in the future to source any new R script that we might save in the R folder. Where Targets will look for sourcing any R script from. 

```{r}
source_all("R")
```

## Data all files combined

We check the final data frame created in the pipeline after merging all three Attendances files

```{r}
tar_read(one_two_combined) 
```

## Type I Accident and Emergency Attendances plot 

This section includes the plot for Type I A&E Attendances created in the pipeline

```{r}
tar_read(plot_1_report)
```

## Type II Accident and Emergency Attendances plot 
