---
title: "A&E Attendances"
author: "PLR"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
if (!require(pacman)){
  install.packages("pacman",dependencies = TRUE)
  library(pacman)
}
pacman::p_load(tidyverse, markdown, janitor, here,gt)
```

## 1. Load required data 
  
We start by loading required A&E Attendances Type I, Type II and Type III data from NHS England website. Link to the data:. Downloaded file "Monthly A&E Time Series March 2024 (XLS,403K)"<https://www.england.nhs.uk/statistics/wp-content/uploads/sites/2/2024/04/Monthly-AE-Time-Series-March-2024.xls>

Use **echo=TRUE** to display both the *code* and *output* of the R code chunk, as in the example below: 

```{r import_AE_data, echo=TRUE, warning=FALSE,message=FALSE}
# Read csv file
AE_data <- read_csv(here("data","AE_Attendances_2010_2024.csv"), 
                                        col_types = cols(...1 = col_skip(), Period = col_date(format = "%d/%m/%Y"), 
                                        Type1_ATT = col_integer(), Type2_ATT = col_integer(), 
                                        Type3_ATT = col_integer())) %>% 
                              clean_names()
head(AE_data)
```

## 2. Reference period  

Today's date is:**`r format(today(),"%d %B")`**. This report was published on the week starting on  **`r format(today()-2,"%d %B %Y")`**. It covers A&E Attendances data from **`r format(min(AE_data$period, na.rm=T),"%d %B %Y")`** as first month in the existing input file to **`r format(max(AE_data$period, na.rm=T),"%d %B 
%Y")`** as the latest available data point.

## 3. Summary Type I, TypeII and TypeIII Attendances

This is a summary of Type I, Type II and Type III attendances figures for the **`r format(min(AE_data$period, na.rm=T),"%d %B %Y")`** to **`r format(max(AE_data$period, na.rm=T),"%d %B 
%Y")`** period: 

```{r Type_1_summary, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Type I figures

Min_Type_I_data <- AE_data %>% select (period, type1_att) %>% filter(type1_att == min(type1_att))
Min_Type_I_data

Max_Type_I_data <- AE_data %>% select (period, type1_att) %>% filter(type1_att == max(type1_att))
Max_Type_I_data


Avg_Type_I_data <- AE_data %>% select (period, type1_att) %>% mutate(AVG_type1_att = mean(type1_att)) %>% filter(period == max(period))
Avg_Type_I_data
```

- For the **`r format(Min_Type_I_data$period,"%d %B %Y")`** to **`r format(Max_Type_I_data$period,"%d %B %Y")`** period, the *lowest* value of AE Type I attendances was **`r prettyNum(Min_Type_I_data$type1_att, big.mark=",")`**. The *highest* value was **`r prettyNum(Max_Type_I_data$type1_att, big.mark=",")`** and the average value for Type I Attendances was
**`r prettyNum(Avg_Type_I_data$type1_att, big.mark=",")`**

```{r Type_2_summary, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Type I figures

Min_Type_II_data <- AE_data %>% select (period, type2_att) %>% filter(type2_att == min(type2_att))
Min_Type_II_data

Max_Type_II_data <- AE_data %>% select (period, type2_att) %>% filter(type2_att == max(type2_att))
Max_Type_II_data


Avg_Type_II_data <- AE_data %>% select (period, type2_att) %>% mutate(AVG_type2_att = mean(type2_att)) %>% filter(period == max(period))
Avg_Type_II_data
```

- The *lowest* value of AE Type II attendances was **`r prettyNum(Min_Type_II_data$type2_att, big.mark=",")`**. The *highest* value was **`r prettyNum(Max_Type_II_data$type2_att, big.mark=",")`** and the **average** value for Type II Attendances was
**`r prettyNum(Avg_Type_II_data$type2_att, big.mark=",")`**

Formatted number **`r prettyNum(Min_Type_I_data$type1_att, big.mark=",")`**

## 4. Year 2023/24 Type I, TypeII, Type III table details 

The table below displays the latest attendances figures by month. 

```{r gt_table,echo=TRUE, warning=FALSE,message=FALSE }

names(AE_data)

AE_data_2013_gt <- AE_data %>% select(Date = period, 
                                 TypeI = type1_att, 
                                 Type2 = type2_att,
                                 Type3 = type3_att) %>% 
                          filter(Date >= '2023-01-01')
# Build a gt table from the previous AE data 
# Formatting numbers with thousands separator, adding title and subtitle
library(gt)
AE_table <-gt(AE_data_2013_gt) 

# Att title to table
AE_table_thousands <- AE_table %>% 
  tab_header(
    title = md("**A&E Attendances in England. 2023 period**") ,
    subtitle = md("By type(*Type I*,*Type II*,*Type III*)")
  ) %>% 
  tab_source_note(
    source_note = "Source: NHE England A&E Attendances and Emergency Admissions data"
  ) %>% 
  tab_source_note(
    source_note = md("*England Time Series monthly data*")
  ) %>% 
  fmt_number( columns = c(TypeI, Type2, Type3), sep_mark = ",", decimals = 0)

AE_table_thousands
```

## 5.Area chart displaying Type I, TypeII, Type III Attendances by Type

This final section describe A&E Attendances by type as an area chart, to visualize the contribution of each A&E Attendances type to the overal total number of Attendances

I will mimic the chart displaying different categories by colours displayed in this website: <https://epirhandbook.com/en/reports-with-r-markdown.html>

```{r Time chart data prep, echo=TRUE}
# First I need to re-shape this data set from wide to long format 

AE_data_all <- AE_data %>% select(Date = period, 
                                 TypeI = type1_att, 
                                 Type2 = type2_att,
                                 Type3 = type3_att) 
# Pivot longer
AE_data_long <- AE_data %>% 
                  select(Date = period, 
                                 TypeI = type1_att, 
                                 Type2 = type2_att,
                                 Type3 = type3_att) %>% 
                  pivot_longer(names_to = "Type",
                               cols = 2:ncol(AE_data))
AE_data_long
```
Then we can plot the data by Attendances type. 

- How to control the amount of break and minor breaks to display with *date_breaks* and *date_minor_breaks*: <https://r-graph-gallery.com/279-plotting-time-series-with-ggplot2.html>
- Also this is an example of a Stacked area chart: <https://r-graph-gallery.com/136-stacked-area-chart.html>


```{r Time chart by Type , echo=TRUE, warning=FALSE,message=FALSE, out.height="100%", out.width="110%"}
library(viridis)
library(hrbrthemes)

ggplot(AE_data_long, aes(x = Date, y = value, fill = Type)) +
      labs(title = "A&E Attendances by type. 2010-2024 period",
         subtitle = "TypeI, TypeII, TypeIII A&E Attendances by month",
         x = "Period", y = "Value" ) +
  geom_area(alpha = 0.6, size = .5, colour = "white") +
  scale_y_continuous(breaks = seq(0, 6000000, by = 300000)) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)) +
   scale_fill_viridis(discrete = T) +
 theme_ipsum() 

```
